---
title: "Class 1 Homework"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the following libraries:

```{r, echo = T, message=F, warning=F, fig.width=5, fig.height=4, fig.align="center"}
library(tidyverse)  
library(kableExtra)
library(gcookbook)

```

Get heightweight data *(built in with the gcookbook package)*, and put it in a dataframe *(the gcookbook:: syntax tell r which package the function or data you want is coming from. This normally is not necessary, unless packages have functions or data with the same name)*:

```{r, echo = T, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

dfHW <- heightweight

head(dfHW)

```
Let's take a quick look at the distribution of female and male observations. Use count to find the totals:

```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

count(dfHW, sex)

```
Now, Create a scatter plot of Weight vs Age in Years:

```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p = ggplot(dfHW, aes(ageYear, weightLb)) + geom_point()
p


```

Change the scatter plot to show sex in different colors:

```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p = ggplot(dfHW, aes(ageYear, weightLb, color = sex)) + geom_point()
p

```

What does that tell us? Is that a little hard to see?

Let's make it easier to see. Create a new dataframe, and group the data by sex and ageYear, Then summarise the means of heightIn. It should look like the following: 

```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

SummaryHW = dfHW %>% group_by(sex, ageYear) %>% summarise(AvgWeight = mean(weightLb))

SummaryHW

```

Now plot that and see what it looks like. And add a geom_smooth regression line to see if we can pick up a trend.

```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p = ggplot(SummaryHW, aes(ageYear, AvgWeight, color = sex)) + geom_point()

p


```

A little easier to see tendency. Do you think there's a tedency? Add a geom_smooth() *(use se = F)*


```{r, echo = F, comment = NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p = p + geom_smooth(se = F)

p


```
