---
title: "Class_4_Notes"
output: pdf_document
---

## JOINS-JOINS-JOINS  


### Getting data from Files and Tidying in the Verse

Recall from your homework - class 3, we used tidyverse inner_join:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

library(DBI)

con1 <- dbConnect(odbc::odbc(), Driver = "{SQL Server}", Server = "analyticslab.database.windows.net", 
    Port = "1433", Database = "A3LProjects", UID = "Student", 
    PWD = "Acct7397")

locationString = "C:/Users/ellen/OneDrive/Documents/UH/Spring 2020/DA1/Section 1/Class 4/"

library(kableExtra)

```

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=5, fig.height=3, fig.align="center"}

library(tidyverse)
library(stringr)
library(lubridate)

OrdDetail = read_csv(str_c(locationString,"SalesOrderDetail.csv"))
Prod = read_csv(str_c(locationString,"Product.csv"))


ProdSales = OrdDetail %>% 
  inner_join(Prod, by = "ProductID") %>%
  group_by(ProductID, Name) %>%
  summarise(TotalSales = round(sum(LineTotal),2)) %>%
  arrange(ProductID)


```



```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(ProdSales$TotalSales)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(ProdSales)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)


```

### Getting Data from the Server, and Tidying in the Verse

Data doesn't always come from text files. MOST data in Enterprises are stored in SQL Servers *(ALL of the ERP, and >90% of Line of Business Applications data is in SQL databases)*.

Lets get data straight off the server:
```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}


# Select Data from Sales Order Detail
OrdDetailSQL <-  dbGetQuery(con1,"
SELECT
[SalesLT].[SalesOrderDetail].[SalesOrderID] 
,[SalesLT].[SalesOrderDetail].[SalesOrderDetailID] 
,[SalesLT].[SalesOrderDetail].[OrderQty] 
,[SalesLT].[SalesOrderDetail].[ProductID] 
,[SalesLT].[SalesOrderDetail].[UnitPrice]
,[SalesLT].[SalesOrderDetail].[UnitPriceDiscount]
,[SalesLT].[SalesOrderDetail].[LineTotal]	
FROM
[SalesLT].[SalesOrderDetail]
")

# Get Data From Product

ProdSQL <- dbGetQuery(con1,"
SELECT 
[SalesLT].[Product].[ProductCategoryID]
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name]
,[SalesLT].[Product].[ProductNumber]
,[SalesLT].[Product].[ListPrice]
,[SalesLT].[Product].[StandardCost]
FROM [SalesLT].[Product]
")


ProdSalesSQL = OrdDetail %>% 
  inner_join(Prod, by = "ProductID") %>%
  group_by(ProductID, Name) %>%
  summarise(TotalSales = round(sum(LineTotal),2)) %>%
  arrange(ProductID)


```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(ProdSalesSQL$TotalSales)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(ProdSalesSQL)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)


```


### Getting Data and Tidying on the Server

Using SQL Server to do the work *(Note how Group By works like group_by R, and Order By works like arrange in R)*:

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

ProdSalesSQL2 <-  dbGetQuery(con1,"
SELECT
[SalesLT].[SalesOrderDetail].[ProductID] 
,[SalesLT].[Product].[Name]
,sum([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM
[SalesLT].[SalesOrderDetail]
INNER JOIN
[SalesLT].[Product]
ON
[SalesLT].[SalesOrderDetail].[ProductID] = [SalesLT].[Product].[ProductID]
GROUP BY 
[SalesLT].[SalesOrderDetail].[ProductID] 
,[SalesLT].[Product].[Name]
ORDER BY
[SalesLT].[SalesOrderDetail].[ProductID] 
")

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(ProdSalesSQL2$TotalSales)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(ProdSalesSQL2)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)


```


Also, note that the SQL INNER JOIN is similar to R.  For application purposes, you should plan to do as much tidying in SQL as possible - Enterprise data is usually **HUGE** in scale, and your desktop isn't going to handle it. and neither will *project* level servers and *marts* like Tableau and Alteryx. That said, you'll still need to do tidying in R too. 

### Adding Product Category and Tidying in the Verse

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

# Get Data From Product with Product Category
Prod <- dbGetQuery(con1,"
SELECT 
[SalesLT].[Product].[ProductCategoryID]
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name]
,[SalesLT].[Product].[ProductNumber]
,[SalesLT].[Product].[ListPrice]
,[SalesLT].[Product].[StandardCost]
FROM [SalesLT].[Product]
")

ProdCat <- dbGetQuery(con1,"
SELECT 
[SalesLT].[ProductCategory].[ProductCategoryID]
,[SalesLT].[ProductCategory].[Name]
FROM [SalesLT].[ProductCategory]
")

ProdCatSales = ProdCat %>% 
  rename(CategoryName = Name) %>% 
  # important to understand how dplyr will handle duplicate column names
  inner_join(Prod, by = "ProductCategoryID") %>%
  inner_join(OrdDetail, by = "ProductID") %>%
  group_by(ProductCategoryID, CategoryName) %>%
  summarise(TotalSales = round(sum(LineTotal),2)) %>%
  arrange(ProductCategoryID)

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(ProdCatSales$TotalSales)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(ProdCatSales)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```


Note the *rename* here. R will not allow you to create a dataframe with 2 columns of the same name *(ProductCategory and Product both have "Name" columns)*. Neither will SQL, but R will rename the column for you *(then you can rename later)*. SQL will just give you an error if you want to store the results in a table - or temp table *(not runnning this below, but it will err out - run in beaver if you want to see)*. SQL won't give you an error if you're just running a report:
```{r, echo = T, eval = FALSE, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

ProductCatSalesSQL = dbGetQuery(con1,"
/*

DROP TABLE IF EXISTS ##TempProductQuery
set nocount on

SELECT
 [SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[Name] 
,sum([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
INTO ##TempProductQuery
FROM
[SalesLT].[ProductCategory]
INNER JOIN
[SalesLT].[Product]
ON
[SalesLT].[ProductCategory].[ProductCategoryID] = [SalesLT].[Product].[ProductCategoryID]
INNER JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
GROUP BY
 [SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[Name] 


*/
")

# No output - just error

```

Note: You can create a temp table in SQL to store results and manipulate data. It's similar to putting a query result into your own dataframe - but it lives on the server for the duration of your session. That said - **don't use temp tables in the this class**. I put it here so you'd be aware of the functionality. You won't need them for classwork or exams.

### Getting Data from the Server, and Tidying in the Server

Without using a temp table, the query would look like this:

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

ProdCatSalesSQL = dbGetQuery(con1,"

SELECT
 [SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] AS CategoryName
,[SalesLT].[Product].[Name] 
,sum([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM
[SalesLT].[ProductCategory]
INNER JOIN
[SalesLT].[Product]
ON
[SalesLT].[ProductCategory].[ProductCategoryID] = [SalesLT].[Product].[ProductCategoryID]
INNER JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
GROUP BY
 [SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[Name] 

")

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(ProdCatSalesSQL$TotalSales)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(ProdCatSalesSQL)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```



## LEFT JOINS

What if management says "I want to see the products that didn't sell too - just as interested in those". A left join includes everything on the *Left* side of the join *(or the right side with a RIGHT JOIN - this works the same in R)*:

### Single Left Join

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

AllProdSalesSQL = dbGetQuery(con1,"
SELECT
 [SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] AS Product_Name
,SUM([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM
[SalesLT].[Product]
LEFT JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
GROUP BY 
 [SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] 

")

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(AllProdSalesSQL$TotalSales, na.rm = T)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(AllProdSalesSQL)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```



### LEFT JOIN with INNER JOIN

Now let's create a LEFT JOIN between Product and SalesOrderDetail with an INNER JOIN between 
Product Category and Product:

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}


AllProdCatSalesSQL = dbGetQuery(con1,"

SELECT
 [SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] AS Product_Name
,SUM([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM 
[SalesLT].[ProductCategory]
INNER JOIN
[SalesLT].[Product]
ON
[SalesLT].[ProductCategory].[ProductCategoryID] = [SalesLT].[Product].[ProductCategoryID]
LEFT JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
GROUP BY 
 [SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] 

")

sum(AllProdCatSalesSQL$TotalSales, na.rm = T)

## BE SURE AND USE na.rm = T when summing a colunn with NAs

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(AllProdCatSalesSQL$TotalSales, na.rm = T)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(AllProdCatSalesSQL)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```


### 2 LEFT JOINs

Extending this example, let's create a LEFT JOIN between Product and SalesOrderDetail with an LEFT JOIN between Product Category and Product:
```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

AllProdCatandProdSalesSQL = dbGetQuery(con1,"
SELECT
 [SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] AS Product_Name
,SUM([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM 
[SalesLT].[ProductCategory]
LEFT JOIN
[SalesLT].[Product]
ON
[SalesLT].[ProductCategory].[ProductCategoryID] = [SalesLT].[Product].[ProductCategoryID]
LEFT JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
GROUP BY 
 [SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] 

")

```


```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

x = "Total Sales = "
y = sum(AllProdCatandProdSalesSQL$TotalSales, na.rm = T)

Out <- data.frame(Description = x, Value = y)

knitr::kable(Out) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

knitr::kable(head(AllProdCatandProdSalesSQL)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```


## SQL Filtering 

One last thing for today. SQL has very sophisticated filtering and aggregation functions, but now for the basics: the **WHERE** clause is roughly equivalent to R's "filter". SQL can also filter aggregates using the **HAVING** clause. See below: 

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

FilterExample = dbGetQuery(con1,"
SELECT
[SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] AS Product_Name
,SUM([SalesLT].[SalesOrderDetail].[LineTotal]) AS TotalSales	
FROM 
[SalesLT].[ProductCategory]
LEFT JOIN
[SalesLT].[Product]
ON
[SalesLT].[ProductCategory].[ProductCategoryID] = [SalesLT].[Product].[ProductCategoryID]
LEFT JOIN
[SalesLT].[SalesOrderDetail]
ON
[SalesLT].[Product].[ProductID] = [SalesLT].[SalesOrderDetail].[ProductID]
WHERE 
[SalesLT].[ProductCategory].[ProductCategoryID] = 5
GROUP BY 
[SalesLT].[ProductCategory].[ProductCategoryID] 
,[SalesLT].[ProductCategory].[Name] 
,[SalesLT].[Product].[ProductID] 
,[SalesLT].[Product].[Name] 
HAVING 
SUM([SalesLT].[SalesOrderDetail].[LineTotal])  < 1000.00

")

```


```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

knitr::kable(FilterExample) %>%
  kable_styling(bootstrap_options = "striped", font_size = 9)

```

