---
title: "Class_2_Homework"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Prep and Visualization

Load the following Libraries and read in ProductInvoices.csv

```{r, echo = T, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

library(tidyverse)
library(stringr)
library(lubridate)

# setwd("/home/ellen/Documents/Spring2020/DA1/Section I/class 2/Class 2 Data Files/")

setwd("C:/Users/ellen/OneDrive/Documents/Spring 2020/DA1/Section 1/Class 2/Class 2 Data Files")

dfProductInvoices = read_csv("ProductInvoices.csv")

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

dfProductInvoices = dfProductInvoices %>% mutate(Date2 = mdy(Date))

dfProductInvoices = dfProductInvoices %>% mutate(Date = Date2) %>% select(-Date2) 


dfProductInvoicesLong = dfProductInvoices %>% 
  pivot_longer(
  cols = 2:ncol(dfProductInvoices),
  names_to = "Product",
  values_to = "Sales",
  values_drop_na = TRUE
)

```

Create a bar plot to visualize sales by date, similar to the following:
```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p1 = ggplot(dfProductInvoicesLong, aes(Date, Sales)) + geom_bar(stat = "identity")
p1

```

Group by week and summarize sales. Then plot showing products in colors *(you can use show.legend = FALSE in the geom_bar function to save a huge display of products)*:
```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

dfProductInvoicesSummary = dfProductInvoicesLong %>% group_by(Week = week(Date), Product) %>% summarise(WkSales = sum(Sales))

p1 = ggplot(dfProductInvoicesSummary, aes(Week, WkSales, fill = Product)) + geom_bar(stat = "identity", show.legend = FALSE) 
p1

```

Rather than stacking products, show product sales separately *(position = "dodge")*

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p1 = ggplot(dfProductInvoicesSummary, aes(Week, WkSales, fill = Product)) + geom_bar(stat = "identity", position = "dodge", show.legend = FALSE)
p1

```

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

dfProductInvoicesSummaryWide <- dfProductInvoicesSummary %>% 
  pivot_wider(names_from = "Product", values_from = "WkSales") 

Forecast = dfProductInvoicesSummaryWide %>% ungroup() %>% filter(Week == 1)  %>% mutate(Week = 4) %>%
  bind_rows(dfProductInvoicesSummaryWide) %>% pivot_longer(
   cols = 2:ncol(dfProductInvoicesSummaryWide),
   names_to = "Product",
   values_to = "Sales",
   values_drop_na = TRUE
 )


```

### Forecast

Now, add a week 4 to the data, with the same sales as week 1. And show that in a stacked bar:

```{r, echo = F, comment=NA, message=F, warning=F, fig.width=4, fig.height=3, fig.align="center"}

p1 = ggplot(Forecast, aes(Week, Sales, fill = Product)) + geom_bar(stat = "identity", show.legend = FALSE) 
p1


```
